Competing Bandits:\\ The Perils of Exploration under Competition \thanks{This is a merged and final version of two conference papers, \citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19}, with a unified and streamlined presentation.}

### 2020-07-11 journal paper/content/sec-intro.tex
Some edits / comments in "network effects" xhdr.

\xhdr{Economic interpretation: network effects \asedit{of data}.}
Our model speaks to the ongoing debate on the role that data plays in the digital economy. One fundamental question in this debate is whether data can serve a similar role as traditional network effects, whereby competition \emph{for} the market is more important than competition \emph{in} the market, due to the winner-take-all nature of \asedit{the digital economy} \citep{Rysman09, jullien2019economics}. 
The ``death spiral" effect that we find has a similar flavor: 
\asdelete{as this as}
once a firm engages in exploration it gets completely starved of users. We further find that a small data advantage for one firm gets amplified under competition and leads to that firm taking the entire market. A distinctive feature of our approach is that we explicitly model the learning problem of the firms and consider them deploying algorithms for solving this problem.  Thus, we do not explicitly model \asedit{the} network effects, but they arise endogenously from our setup.

\ascomment{Guy: could you pls clarify the diff between "for the market" and "in the market". Is this the distinction between "competing for market share" and "competing for profit"? If so, I understand what you mean by "for" the market; but why "in" the market? Perhaps a small edit would make it much clearer?}

Our results highlight that understanding the performance of learning algorithms in isolation does not necessarily translate to understanding their impact in competition, precisely due to the fact that competition leads to the endogenous generation of data observed by the firms. Approaches such as \citet{bajari2018impact} argue that the diminishing returns to scale and scope of data in isolation \asedit{mitigate} such data feedback loops,
%as non-existent
\asedit{but} ignore the differences induced by learning in isolation versus under competition. Furthermore, explicitly incorporating the interaction between learning technology and data creation allows us to speak on how data advantages are characterized and amplified not only by data quantity, but also the increased data quality gathered by better learning algorithms.

@@

>>>>

\xhdr{Economic interpretation: network effects of data.}
Our model speaks to the ongoing debate on the role that data plays in the digital economy. One fundamental question in this debate is whether data can serve a similar role as traditional ``network effects", \asedit{creating scenarios when only one firm can function in the market \citep{Rysman09, jullien2019economics}.}
%whereby, when these effects are present, in many cases only one firm can function in the market, leading to competition \emph{for} the market being more important than competition \emph{in} the market \citep{Rysman09, jullien2019economics}.
\asedit{The death spiral/amplification effects mentioned above have a similar flavor as network effects: a relatively small amount of exploration (resp., data advantage)  gets amplified under competition and causes the firm to be starved out of users (resp., take over most of the market).}
%\gaedit{We further find that a small data advantage for one firm gets amplified under competition and leads to that firm taking the entire market, showing that data can provide a similar incumbency advantage as those provided by traditional network effects and can serve as a barrier to entry in online markets.} 
A distinctive feature of our approach is that we explicitly model the learning problem of the firms and consider them deploying algorithms for solving this problem.  Thus, we do not explicitly model the network effects, but they arise endogenously from our setup.

Our results highlight that understanding the performance of learning algorithms in isolation does not necessarily translate to understanding their impact in competition, precisely due to the fact that competition leads to the endogenous generation of data observed by the firms. Approaches such as \citet{lambrecht2015can, bajari2018impact} argue that the diminishing returns to scale and scope of data in isolation mitigate such data feedback loops,
%as non-existent
but ignore the differences induced by learning in isolation versus under competition. Furthermore, explicitly incorporating the interaction between learning technology and data creation allows us to speak on how data advantages are characterized and amplified not only by data quantity, but also the increased data quality gathered by better learning algorithms.

@@

\gacomment{I added back the \cite{lambrecht2015can} reference, but \cite{varian2018artificial} is a more informal discussion of these issues and not essential to include. However, even the \cite{lambrecht2015can} is relatively informal discussion of the same issues. Part of the problem that makes this tricky is that the literature on network effects is a well-developed and mature literature, but the question of whether data plays a similar role primarily started as more of a policy discussion, of which there's only starting to be some serious theoretical and empirical work on the issue.}
